---
title: "Stairtek Price Variation"
author: "Jeevan Sai Reddy, Bharath Vattikuti, Jeffrey Griffiths"
date: "January 11, 2019"
output: 
  ioslides_presentation:
    font_adjustment: -1
---


<style type="text/css">
  
  body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
    font-size: 8px;
}
h1.title {
  font-size: 120px;
  color: DarkRed;
}
h1 { /* Header 1 */
    font-size: 120px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 32px;
  color: DarkBlue;
  line-height: 30px;
}
h3 { /* Header 3 */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}

code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>


<style type="text/css">
body p {
  color: #000000;
  font-size: 20px;
}
</style>

<style>
table {
  background-color: white !important;
  color: black !important;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r echo = FALSE, warning=FALSE, message=FALSE}

# Set working directory
setwd("C:\\Users\\jeeva\\Documents\\BANA\\Career Management\\Amend Usecase\\")
library(tidyverse)
library(lubridate)
library(ggplot2)
library(gridExtra)
library(knitr)
library(scales)
library(plotly)

setwd("C:\\Users\\jeeva\\Documents\\BANA\\Career Management\\Amend Usecase\\")
amend_data <- read.csv("Amend_use_case.csv", header = TRUE, stringsAsFactors = FALSE)

amend_data$unit_price <- ifelse(amend_data$Last.Unit.Cost == 0, amend_data$Original.Unit.Cost, 
                                amend_data$Last.Unit.Cost)
amend_data$final_cost <- amend_data$Received.Quantity * amend_data$unit_price

#hist(amend_data$final_cost)
#summary(amend_data)
amend_data$order_date2 <- as.Date(amend_data$Order.Date, format = "%m/%d/%Y")
amend_data$month <- as.Date(cut(amend_data$order_date2,breaks = "month"))

```
}

## Quantity ordered and Amount spent have high variability{.smaller}

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 8}
# Amount spent by month
spend_plot <- amend_data %>% 
  group_by(month) %>%
  summarise(total_cost = sum(final_cost, na.rm = TRUE)) %>%
  ggplot(aes(x = month, y = total_cost, group = 1)) +
  geom_line() +
  geom_point() + 
  scale_x_date(labels = date_format("%Y-%m"),breaks = "2 month") +
  scale_y_continuous(name = "Total Spent", labels = scales::dollar) +
  labs(x = "Month", title = "Amount spent by month")

# Quantity by month
quantity_plot <- amend_data %>% 
  group_by(month) %>%
  summarise(total_quan = sum(Received.Quantity, na.rm = TRUE)) %>%
  ggplot(aes(x = month, y = total_quan, group = 1)) +
  geom_line() +
  geom_point() + 
  scale_x_date(labels = date_format("%Y-%m"),breaks = "2 month") +
  scale_y_continuous(name = "Total Quantity", labels = scales::comma) +
  labs(x = "Month", title = "Quantity by month")

grid.arrange(spend_plot, quantity_plot, ncol = 1)

```

Overall, quantity ordered and amount have high variability. Also, quantity and amount aren't correlated with each other


## Which lead to high variability in Price Per Unit{.smaller}
```{r echo = FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 8}

# Amount spent by month
ppu_plot <- amend_data %>% 
  group_by(month) %>%
  summarise(price_per_unit = sum(final_cost, na.rm = TRUE) / sum(Received.Quantity, na.rm = TRUE)) %>%
  ggplot(aes(x = month, y = price_per_unit, group = 1)) +
  geom_line() +
  geom_point() + 
  scale_x_date(labels = date_format("%Y-%m"),breaks = "2 month") +
  scale_y_continuous(name = "Price Per Unit", labels = scales::dollar) +
  labs(x = "Month", title = "Price Per Unit across months")

ggplotly(ppu_plot)

```

<span style="color:black">
<br>
As expected, Price Per Unit also has high variability across months
</span>


## 80% of amount was spent on top 3 item codes{.smaller}

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 8}

# To Understand quantity and amount spent by item code
amend_data_item_code <- amend_data %>% 
  group_by(Item.Code.Acronym) %>%
  summarise(total_spent = sum(final_cost, na.rm = TRUE), 
            total_quantity = sum(Received.Quantity, na.rm = TRUE)) %>%
arrange(desc(total_spent)) 

# Calculating spent percentage
amend_data_item_code$spent_percent <- round((amend_data_item_code$total_spent / 
                                         sum(amend_data_item_code$total_spent))*100,2)

amend_data_item_code$quan_percent <- round((amend_data_item_code$total_quantity / 
                                              sum(amend_data_item_code$total_quantity))*100,2)

format_money  <- function(x, ...) {
  paste0("$", formatC(as.numeric(x), format="f", digits=0, big.mark=","))
}

format_number  <- function(x, ...) {
  formatC(as.numeric(x), format="f", digits=0, big.mark=",")
}

format_percent <- function(x, ...) {
  paste0(formatC(x, format = "f", digits = 0, ...), "%")
}

amend_data_item_code$total_spent <- format_money(amend_data_item_code$total_spent)
amend_data_item_code$total_quantity <- format_number(amend_data_item_code$total_quantity)
amend_data_item_code$spent_percent <- format_percent(amend_data_item_code$spent_percent)
amend_data_item_code$quan_percent <- format_percent(amend_data_item_code$quan_percent)

names(amend_data_item_code) <- c("Item_Code", "Total_Spent", "Total_Quantity",
                                 "Spent_Percent", "Quan_Percent")

amend_data_item_code <- amend_data_item_code[c(1:3),]

kable(amend_data_item_code, caption = "Summary of spent and quantity by item code")

```

<span style="color:black">
<br>
Around 80% of spend is on top 3 items and these top 3 items correspond to approximately 20% of total quantity ordered
</span>

## Even at item code level, there is high variability in Price Per Unit{.smaller}

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 8}

# Aggregating at Item code and month level
amend_df_aggregated <- amend_data %>% 
  group_by(month, item_code = Item.Code.Acronym) %>%
  summarise(total_quan = sum(Received.Quantity, na.rm = TRUE), 
            total_price = sum(final_cost, na.rm = TRUE)) %>%
  mutate(price_per_unit = total_price / total_quan)

# Filtering only top 3 item codes
top3_item_codes <- amend_df_aggregated %>%
  group_by(item_code) %>%
  summarise(total_price = sum(total_price)) %>%
  arrange(desc(total_price)) %>%
  top_n(3) %>%
  select(item_code)

top3_item_codes <- as.vector(top3_item_codes$item_code)

top3_unit_price_plot <- amend_df_aggregated %>%
  filter(item_code %in% top3_item_codes) %>%
  ggplot(aes(x = month, y = price_per_unit, colour = item_code, group = item_code)) +
  geom_line() + 
  geom_point() + 
  scale_x_date(labels = date_format("%Y-%m"),breaks = "2 month") +
  scale_y_continuous(name = "Price Per Unit", labels = scales::dollar) +
  scale_fill_discrete(name = "Item Code") +
  labs(x = "Month", title = "Price Per Unit for top 3 item codes")

ggplotly(top3_unit_price_plot)
```

<span style="color:black">
<br>
Price per unit has high variation for BOT and BOX. Variation is considerably less for RIS
</span>

##Splitting further to understand variation at Species level{.smaller}

<span style="color:black">
This table shows number of species where coefficeint of variation at species level is greater or lesser than overall coefficient of variation (i.e. at Item code level)
</span>

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 8}

amend_data$month_yr <- format(amend_data$order_date2,"%Y-%m")
# Finding unit price for BOT at species level
top5_species_unit_price_bot <- amend_data %>%
  filter(Item.Code.Acronym %in% c('BOT')) %>%
  group_by(month_yr, Species, Item.Code.Acronym) %>%
  summarise(tot_price = sum(final_cost, na.rm = TRUE), tot_quan = sum(Received.Quantity, na.rm = TRUE)) %>%
  mutate(unit_price = tot_price / tot_quan)

# Finding unit price for BOX at species level
top5_species_unit_price_box <- amend_data %>%
  filter(Item.Code.Acronym %in% c('BOX')) %>%
  group_by(month_yr, Species, Item.Code.Acronym) %>%
  summarise(tot_price = sum(final_cost, na.rm = TRUE), tot_quan = sum(Received.Quantity, na.rm = TRUE)) %>%
  mutate(unit_price = tot_price / tot_quan)

# Finding unit price for RIS at species level
top5_species_unit_price_ris <- amend_data %>%
  filter(Item.Code.Acronym %in% c('RIS')) %>%
  group_by(month_yr, Species, Item.Code.Acronym) %>%
  summarise(tot_price = sum(final_cost, na.rm = TRUE), tot_quan = sum(Received.Quantity, na.rm = TRUE)) %>%
  mutate(unit_price = tot_price / tot_quan)

# Filtering out NAs
top5_species_unit_price_bot <- top5_species_unit_price_bot[complete.cases(top5_species_unit_price_bot),]
top5_species_unit_price_box <- top5_species_unit_price_box[complete.cases(top5_species_unit_price_box),]
top5_species_unit_price_ris <- top5_species_unit_price_ris[complete.cases(top5_species_unit_price_ris),]

# Find coefficeint of variation at species level for BOT
df_bot <- top5_species_unit_price_bot %>%
  group_by(Species) %>%
  summarise(coeff_var = sd(unit_price) / mean(unit_price))
df_bot$item_code <- 'BOT'

# Find coefficeint of variation at species level for BOX
df_box <- top5_species_unit_price_box %>%
  group_by(Species) %>%
  summarise(coeff_var = sd(unit_price) / mean(unit_price))
df_box$item_code <- 'BOX'

# Find coefficeint of variation at species level for RIS
df_ris <- top5_species_unit_price_ris %>%
  group_by(Species) %>%
  summarise(coeff_var = sd(unit_price) / mean(unit_price))
df_ris$item_code <- 'RIS'

# Combining in to one CV df
cv_df_combined <- rbind(df_bot, df_box, df_ris)

# Creating one data frame to calculate CV at Item Code level
cv_df_all_species <- amend_data %>%
  filter(Item.Code.Acronym %in% c('BOT', 'BOX', 'RIS')) %>%
  group_by(month_yr, Item.Code.Acronym) %>%
  summarise(tot_price = sum(final_cost, na.rm = TRUE), tot_quan = sum(Revised.Quantity, na.rm = TRUE))

# Filtering out NA values  
cv_df_all_species <- cv_df_all_species[which(cv_df_all_species$tot_price > 0 &
                                               cv_df_all_species$tot_quan > 0),]

# Calculating unit price
cv_df_all_species$unit_price <- cv_df_all_species$tot_price / cv_df_all_species$tot_quan

# 2018-11 has very low unit price. Need to understand data quality here
cv_df_all_species <- cv_df_all_species[cv_df_all_species$month_yr != '2018-11', ]

# Finding CV at item code level
cv_df_all_species_v2 <- cv_df_all_species %>%
  group_by(Item.Code.Acronym) %>%
  summarise(cv_item = sd(unit_price) / mean(unit_price))

# Filtering out NA values
cv_df_combined <- cv_df_combined[complete.cases(cv_df_combined), ]

colnames(cv_df_all_species_v2) <- c("item_code", "coeff_var_overall")

# Obtaining one final data frame where we get CV at species and CV for Item code
cv_df_combined_v2 <- merge(x = cv_df_combined, y = cv_df_all_species_v2, by = "item_code")

# Splitting above data frame in to two
# 1. where cv of species is greater than overall CV i.e. item code
# 2. where cv of species is less than overall CV i.e. item code
cv_df_species_greater <- cv_df_combined_v2 %>%
  filter(coeff_var > coeff_var_overall)

cv_df_species_lesser <- cv_df_combined_v2 %>%
  filter(coeff_var <= coeff_var_overall)

# Creating one data frame which is a combination of species with cv greater than 
# overall and less than overall
cv_summary_df <- merge(cv_df_species_greater %>%
                       group_by(item_code) %>%
                       summarise(num_species_greater = n_distinct(Species)),
                       cv_df_species_lesser %>%
                       group_by(item_code) %>%
                       summarise(num_species_lesser = n_distinct(Species)),
                       by = "item_code", all = TRUE)

cv_summary_df[is.na(cv_summary_df)] <- 0

cv_summary_df$greater_percent <- round((cv_summary_df$num_species_greater / (cv_summary_df$num_species_greater + cv_summary_df$num_species_lesser) )*100,2)

cv_summary_df$greater_percent <- format_percent(cv_summary_df$greater_percent)

names(cv_summary_df) <- c("Item_Code",
                          "Num_Species_Greater",
                          "Num_Species_Lesser",
                          "Greater_Percent")

kable(cv_summary_df, caption = "Summary of coefficient of variation")

```

<span style="color:black">
<br>

* For all species in BOX, variation in price is less than overall price. Hence variation here could be because of order strategy
 
* For BOT around 53% species have variation greater than overall and for RIS around 41% have variation greater than overall
</span>


## Around 50% of species from RTR have higher variability{.smaller}

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 8}

cv_df_species_greater_manuf <- 
  merge(
    amend_data %>%
      group_by(item_code = Item.Code.Acronym, Species, Vendor.Name) %>%
      summarise(tot_sales = sum(final_cost, na.rm = TRUE),
                total_quan = sum(Received.Quantity, na.rm = TRUE)),
      cv_df_species_greater,
    by = c("item_code", "Species")
  )

cv_df_species_lesser_manuf <- 
  merge(
    amend_data %>%
      group_by(item_code = Item.Code.Acronym, Species, Vendor.Name) %>%
      summarise(tot_sales = sum(final_cost, na.rm = TRUE),
                total_quan = sum(Received.Quantity, na.rm = TRUE)),
    cv_df_species_lesser,
    by = c("item_code", "Species")
  )

# More Analysis on manufacturer

cv_summary_df <- merge(
x = cv_df_species_greater_manuf %>%
  group_by(Vendor.Name) %>%
  summarise(tot_species_great = n_distinct(Species), 
            tot_sales_great = sum(tot_sales),
            total_quan_great = sum(total_quan)) %>%
  mutate(price_per_quan_great = tot_sales_great / total_quan_great),

y = cv_df_species_lesser_manuf %>%
  group_by(Vendor.Name) %>%
  summarise(tot_species_less = n_distinct(Species), 
            tot_sales_less = sum(tot_sales),
            total_quan_less = sum(total_quan)) %>%
  mutate(price_per_quan_less = tot_sales_less / total_quan_less),

by = "Vendor.Name",
all.x = TRUE,
all.y = TRUE
)

cv_summary_df[is.na(cv_summary_df)] <- 0

cv_summary_df$tot_species_great_percent <- (cv_summary_df$tot_species_great / 
                                         (cv_summary_df$tot_species_great + 
                                            cv_summary_df$tot_species_less))*100




cv_summary_df2 <- cv_summary_df[,c("Vendor.Name", "tot_species_great", 
                                  "tot_species_less", "tot_species_great_percent",
                                  "price_per_quan_great", 
                                  "price_per_quan_less")]

cv_summary_df2$tot_species_great <- format_number(cv_summary_df2$tot_species_great)
cv_summary_df2$tot_species_less <- format_number(cv_summary_df2$tot_species_less)
cv_summary_df2$tot_species_great_percent <- format_percent(cv_summary_df2$tot_species_great_percent)
cv_summary_df2$price_per_quan_great <- format_money(cv_summary_df2$price_per_quan_great)
cv_summary_df2$price_per_quan_less <- format_money(cv_summary_df2$price_per_quan_less)


names(cv_summary_df2) <- c("Vendor_Name",
                           "High_CV_Species",	
                           "Low_CV_Species"	,
                           "%High_CV_Species",	
                           "High_CV_PPQ"	,
                           "Low_CV_PPQ")

kable(cv_summary_df2, caption = "Variation across vendors")
```

<span style="color:black">
<br>

* Around 50% of species ordered from RTR have higher variability and same number is 20% from CFFCO

* Price Per Unit is higher in high variability Species indicating indicating higher the unit price, higher the variation
</span>


## Built various Machine Learning Models to predict price {.smaller}

**Below are important metrics considered**
<span style="color:black">

 * Quantity
 * Item Code
 * Species
 * Vendor
 * Time taken to close order
 
</span>  

**Below are various Machine Learning Models considered**

 * Linear Regression
 * Decison Trees
 * Random Forest
 * Gradient Boosting Machines

</span>


## Able to predict expected price with average deviation of 15%

<span style="color:black">

* Built around 4k models to choose best performing model
* Built an automated dashboard which gives expected price as output
* Dashboard demo

**Next Steps**

* Settingup Dashboard for production
* Testing Model
* Model in to prodcution

</span>